---
title: "p8105_hw3_smm2350"
author: "Shaina Mackin" 
date: "2021-10-20"
output: github_document
---

# Homework 3

To begin, I will load the tidyverse and instacart data, as well as set knitr options.

```{r libraries, echo = FALSE, message = FALSE}
library(tidyverse)
library(p8105.datasets)
data("instacart")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

### data exploration

I will write a short description of the dataset, noting the size and structure of the data, describing key variables, and giving illustrative examples of observations:

The instacart dataset contains online grocery shopping data from 2017, with `r nrow(instacart)` observations, or products from orders, and `r ncol(instacart)` variables, for a total of `r (nrow(instacart)*ncol(instacart))` data points. Key variables include `order_id`, `product_id`, `user_id`, whether items were `reordered`, `order_hour_of_day` and `days_since_prior_order` (ranging from `r min(pull(instacart, days_since_prior_order))` to `r max(pull(instacart, days_since_prior_order))` ), as well as `product_name` and associated `aisle` and `department`.

```{r}

```

